%! TEX root=../main.tex
\subsection{Abelian extension}

\begin{definition}
  $L/K$ is called an abelian extension if $L/K$ is Galois and $\Gal(L/K)$ is abelian.
\end{definition}

\begin{example}
  For an extension $\Fb_{q^n} / \Fb_q$ of a finite field, $\Fb_{q^n}$ is a splitting field of $x^{q^n}-x$
  over $\Fb_p$, so $\Fb_{q^n} / \Fb_q$ is Galois by theorem~\ref{thm:splitting-field-iff-finite-normal}.
  By theorem~\ref{thm:aut-of-finite-field}, we know that $\Gal(F_{q^n} / F_q) = \langle \sigma_q \rangle$
  is a cyclic group.
\end{example}

\begin{definition} \hfill
  \begin{itemize}
    \item The cyclotomic field $\Qb(\zeta_n)$ is the splitting field of $x^n - 1$ over $\Qb$.
    \item $\zeta$ is called an nth root of unity if $\zeta^n = 1$. $\mathcal{U} = \langle \zeta \rangle$
      is the multiplicative group of nth roots of unity.
    \item $\zeta_n$ is called a primitive nth root of unity if $\zeta^n = 1$ but $\zeta^m \ne 1, \, \forall 0 < m < n$.
    \item The nth cyclotomic polynomial is defined as
      \[ \Phi_n \triangleq \prod (x - \zeta) = \prod_{\gcd(k, n) = 1} (x - \zeta_n^k) \implies \deg \Phi_n = \varphi(n) \]
  \end{itemize}
\end{definition}

\begin{prop} \hfill
  \begin{itemize}
    \item $x^n - 1 = \prod_{d \mid n} \Phi_d$.
      \begin{proof}
        First, Both sides have no multiple root. Then since $\alpha^n = 1 \iff \ord_{\times}(\alpha) \mid n$,
        we know that two sides has equal roots.
      \end{proof}
    \item $\Phi_n \in \Zb[x]$.
      \begin{proof}
        By induction on $n$. $n = 1$ is trivial.
        Assume that the statement is true for all $k < n$, then since
        \[ x^n - 1 = \Phi_n \prod_{\substack{d \mid n, d < n}} \Phi_d \triangleq \Phi_n \Phi_{< n} \]
        But notice that $\Phi_{<n}$ is monic, so by the long division algorithm, it is easy to
        see that $\Phi_n = (x^n - 1) / \Phi_{<n}$ has all coefficient in $\Zb$.
      \end{proof}
    \item $\Phi_n$ is irreducible.
      \begin{proof}
        Suppose $\Phi_n = f(x) g(x)$ with $f$ irreducible, and both $f, g$ are monic.
        By Gauss lemma, we could assume $f(x), g(x) \in \Zb[x]$.
        Let $\zeta_n$ be a primimtive nth root of unity  which satisfied $f(\zeta_n) = 0$
        and $p$ be a prime with $p \nmid n$.

        Assume that $g(\zeta_n^p) = 0$, $m_{\zeta_n, \Qb} = f \implies f \mid g(x^p)$,
        say $g(x^p) = f(x) h(x)$.
        By the long division algorithm, we know that $h(x) \in \Zb[x]$, since $f(x) \in \Zb[x]$
        and monic.

        In $\Zb / p\Zb[x]$, we have $\bar{g}(x)^p = \bar{g}(x^p) = \bar{f}(x) \bar{h}(x)$,
        which implies $\bar{g}, \bar{f}$ has common root, thus $\bar\Phi_n = \bar{f}\bar{g}$ and
        hence $x^n - \bar{1}$ has a multiple root.
        But $(x^n - \bar{1})' = nx^{n-1} \neq 0$, and $0$ is not a root of $x^n - \bar{1}$,
        which leads to an contradiction.

        So we conclude that $f(\eta_n^p) = 0$ for any $p \mid n$, which could be extend
        and show that $f(\eta_n^k) = 0$ for any $\gcd(k, n) = 1$, thus $f = \Phi_n$.
      \end{proof}
    \item $\Qb(\zeta_n) / \Qb$ is Galois with $[\Qb(\zeta_n): \Qb] = \deg \Phi_n = \varphi(n)$.
    \item $\Gal(\Qb(\zeta_n) / \Qb) \cong \Fb_n^\times$.
  \end{itemize}
\end{prop}

\begin{example}
  Let $n = p$ be a prime.
  \begin{itemize}
    \item $\Gal(\Qb(\zeta_p) / \Qb) = \Fb_p^\times = \Zb / (p-1)\Zb$.
    \item For $H \leq \Gal(\Qb(\zeta_p) / \Qb)$, find $\Qb(\zeta_p)^H$. 
      $\alpha = \sum_{\tau \in H} \tau(\zeta_p) \implies \alpha \in \Qb(\zeta_p)^H$

  \end{itemize}
\end{example}
\begin{example}
  Suppose $n = p_1^{n_1} p_2^{n_2} \cdots p_k^{n_k}$ with $p_1, \dots, p_k$ be distinct primes.
  Define $L_i \triangleq \Qb(\zeta_{p_i^{n_i}})$.
\end{example}

\begin{lemma}
  If $L_1 / K$, $L_2 / K$ are Galois, then $L_1 \cap L_2 / K$, $L_1 L_2 / K$ are Galois and
  \[ \Gal( L_1 L_2 / K) \cong \big\{ (\sigma, \tau) \mid \sigma\big|_{L_1 \cap L_2} = \tau\big|_{L_1 \cap L_2} \big\}
    \leq \Gal(L_1 / K) \times \Gal(L_2 / K) \]
  In particular, if $L_1 \cap L_2 = K$, then $\Gal(L_1 L_2 / K) \cong \Gal(L_1 / K) \times \Gal(L_2 / K)$.

  \begin{proof}
    We know that $L_1 \cap L_2 / K$ is finite, separable.
  \end{proof}
\end{lemma}

Back to our problem, $[L_1 L_2 \cdots L_k: \Qb] = [\Qb(\zeta_n): \Qb] = \varphi(n) = \varphi(p_1^{n_1})
\cdots \varphi(p_k^{n_k}) = [L_1 : \Qb] [L_2 : \Qb] \cdots [L_k : \Qb]$.

\begin{theorem}
  Let $G$ be a finite abelian group. Then exists a subfield $L$ of a cyclotomic field satisfied $\Gal(L / \Qb) \cong G$.
\end{theorem}

\subsection{Kammer extension}
In this section, we assume that $\Char K \nmid n$ and $\eta$ is a primitive $n$th root of unity.

\begin{definition} \hfill
  \begin{itemize}
    \item $L/K$ is called a Kammer extension of exponent $n$ if $\eta \in K$ and $L$ is a splitting field
      of $(x^n - a_1) (x^n - a_2) \dots (x^n - a_k)$ over $K$.
    \item Let $\abs{G} < \intfy$, the exponent $e(G)$ of $G$ is the least positive integer $m$
      satisfied $g^m = 1$ for any $g \in G$.
  \end{itemize}
\end{definition}

\begin{theorem}
  Let $L$ be a splitting field of $x^n - a$ over $K$, then $\Gal(L / K(\eta))$ is cyclic of
  degree dividing $n$. More over $x^n - a$ is irreducible over $K(\eta)$ $\iff$ $[L: K(\eta)] = n$.
\end{theorem}

\begin{definition}
  $L/K$ is called a cyclic extension if $L/K$ is Galois and $\Gal(L/K)$ is cyclic.
\end{definition}

\begin{theorem}
  If $L/K$ is a cyclic extension of degree $n$ and $\eta \in K$, then $L$ is a splitting field of
  some irreducible polynomial $x^n - a$ over $K$.

  \begin{proof}
    Recall a result proved in HW problem: Distinct automorphisms of $L$ are linearly independent over $L$.
    
    Let $\Gal(L/K) = \langle \sigma \rangle$ with $\ord(\sigma) = n$. Then
    \[ \text{id}_L + \eta \sigma + \eta^2 \sigma^2 + \dots + \eta^{n-1} \sigma^{n-1} = 0
      \implies \exists c \in L, \text{ s.t. } \alpha = c + \eta \sigma(c) + \eta^2 \sigma^2(c)
      + \dots + \eta^{n-1} \sigma^{n-1}(c) \neq 0 \]

    Observe that $\sigma(\alpha) = \eta^{-1} \alpha$, so $\alpha \not\in K$.

    Also $\sigma(\alpha) = \eta^{-1}\alpha \in K(\alpha)$, $\sigma\big|_{K(\alpha)} \in \Gal(K(\alpha) / K)$
  \end{proof}
\end{theorem}

\begin{theorem}
  If $L/K$ is Galois which satisfied $\Gal(L/K)$ is abelian of exponent $n$ and $\eta \in K$,
  then $L/K$ is a Kummer extension.

  \begin{proof}
    By induction on $[L: K]$. If $[L: K] = 1$ then $L = K$ so $n = 1$.

    Assume $[L: K] > 1$, then by FTFGAG, $\Gal(L/K) \cong \quot{\Zb}{d_1 \Zb}
    \times \quot{\Zb}{d_2 \Zb} \times $
  \end{proof}
\end{theorem}

\begin{theorem}
  If $L/K$ is a kummer extension of exponent $n$, then $\Gal(L/K)$ is abelian of exponent dividing $n$.
\end{theorem}

\subsubsection{Cubic equations}

\begin{lemma}
  Let $\Char K \neq 2$ and $f(x) \in K[x]$ with $\deg f = n$. Let $L = K(\alpha_1, \dots, \alpha_n)$ be
  a splittig field of $L$ over $K$.

  Define $\delta = \prod_{1 \leq i < j \leq n} (\alpha_i - \alpha_j)$, then $L^{\Gal(L/K) \cap A_n} = K(\delta)$
  and hence $\Gal(L/K) \hookrightarrow S_n$.
\end{lemma}

\begin{prop}
  Let $f(x) = x^3 + px + q$ be irreducible in $K[x]$ and $L$ be a splitting field,
  \begin{itemize}
    \item If $\Gal(L/K) \cong S_3$ then $\sqrt{D} \not\in K$.
    \item If $\Gal(L/K) \cong A_3$ then $\sqrt{D} \in K$.
  \end{itemize}
\end{prop}
